\vspace{60pt}\\ \huge \bfseries Skedge \\ \vspace{10pt} \Large Smarter course scheduling for our University

### 2016-04-14 analytics/npa.tex
submitted

  The results are shown in Figure \ref{fig:navs-browse}. The difference between these results and the initial combined results of \ref{fig:navs-combined} are more substantial than last time, and may defy expectations at first glance. While it may seem intuitive that browsed-for courses should have higher amounts of navigations, note that the graph itself is showing \emph{effectiveness} of such browsing. In this case, it makes sense that a significantly higher proportion of adds/bookmarks occur within zero navigations, because that indicates a high-level search and subsequent adds/bookmarks from that list instead of moving on to another search.

  As expected, and unlike the results of further direct-search analysis, the sections added/bookmarked when browsing within 0 navigations were mostly \emph{main} sections (only 37\% of them were subsections). This makes sense, as users may not be concerned with subsections if the goal of the browse is merely to mark interesting courses. This criterion saw more drastic drops in subsection amounts as well, as the number of navs increased: 5\% subsections for ${navs}=1$, and 4\% subsections for ${navs}=2$.

@@

  Given more time with Skedge Social being live and more forethought in tracking its usage, similar data analysis could be done on the subject of the effectiveness of Skedge in peer-guided course finding. Unfortunately, usage data to answer questions such as ``how many Facebook friends were listed as liking or taking a course when it was added?'', and cannot be retroactively computed due to Skedge's purposeful privacy limitations (Facebook friend data can only be retrieved client-side---no data is stored on the server that could authorize such a data request). Thus, part (c) of hypothesis \#2 cannot be verified.

>>>>

  The results are shown in Figure \ref{fig:navs-browse}. The difference between these results and the initial combined results of Figure \ref{fig:navs-combined} are more substantial than last time, and may defy expectations at first glance. While it may seem intuitive that browsed-for courses should have higher amounts of navigations, note that the graph itself is showing \emph{effectiveness} of such browsing. In this case, it makes sense that a significantly higher proportion of adds/bookmarks occur within zero navigations, because that indicates a high-level search and subsequent adds/bookmarks from that list instead of moving on to another search.

  As expected, and unlike the results of further direct-search analysis, the sections added/bookmarked when browsing within 0 navigations were mostly \emph{main} sections (only 37\% of them were subsections). This makes sense, as users may not be concerned with subsections if the goal of the browse is merely to mark interesting courses. This criterion saw more drastic drops in subsection amounts as well, as the number of navigationss increased: 5\% subsections for ${navs}=1$, and 4\% subsections for ${navs}=2$.

@@

  Given more time with Skedge Social being live and more forethought in tracking its usage, similar data analysis could be done on the subject of the effectiveness of Skedge in peer-guided course finding. Unfortunately, usage data to answer questions such as ``how many Facebook friends were listed as liking or taking a course when it was added?'' was not tracked, and cannot be retroactively computed due to Skedge's purposeful privacy limitations (Facebook friend data can only be retrieved client-side---no data is stored on the server that could authorize such a data request). Thus, part (c) of hypothesis \#2 cannot be verified.

### 2016-04-14 analytics/usage.tex
submitted

It should also be noted that before March, the majority of the userbase as I knew it consisted of seniors, as the majority of people that I knew personally and shared Skedge with belonged to my class. The latest two major events reported in the graph above thus very impressively consist of \emph{very few seniors}, who have no courses left to schedule. March also marked the very first time I publicly advertised Skedge, although not very strongly. It was sent out to all students in the Computer Science mailing list as well as posted to two class-wide Facebook groups.

The usability of and user satisfaction with a tool is hard to measure when it holds a monopoly. Considering that Skedge is an \emph{optional alternative} to CDCS, such consistent usage data (especially with 75\% of sessions being from returning users) undeniably prove its measures of usability and user satisfaction relative to CDCS.

@@

Figure \ref{fig:searchtypes} validates the hypothesis made in Section 2.3 regarding the search interface design, demonstrating that the \emph{department code} and \emph{course number} search modes dominate user searches, accounting for over 85\% of all searches on Skedge. Bringing the nondistinct field for this in CDCS into prominence is thus a design choice that accomodates empirical behavior.

@@

  This already is evidence of an effective search mechanism, but of those empty searches, the vast majority contain typos or are for courses that simply don't exist. The other significant portion of searches include instructor names without the requisite ``taught by'' for SkedgeQL to pick it up as an instructor search, and this issue could easily be fixed.

@@

Making the case for usability superior to CDCS's in the space of mobile support is easy, given that CDCS does not have a mobile-responsive website. But to illustrate the importance of ensuring good user experience on mobile devices, consder that 11.54\% of all Skedge sessions are from mobile devices, which is very substantial. On average, 2.74 pages viewed are viewed and 2 minutes are spent per session on mobile devices.

@@

  \caption[Course blocks in Skedge]{Hovering over a course block in Skedge's side-schedule shows at-a-glance information about it; clicking on it navigates the user to that course}
  \label{fig:searchtypes}

>>>>

It should also be noted that before March, the majority of the userbase as I knew it consisted of seniors, as the majority of people that I knew personally and shared Skedge with belonged in my class year. The latest two major events reported in the graph above thus very impressively consist of \emph{very few seniors}, who have no courses left to schedule. March also marked the very first time I publicly advertised Skedge, although not very strongly. It was sent out to all students in the Computer Science mailing list as well as posted to two class-wide Facebook groups.

The usability of and user satisfaction with a tool is hard to measure when it holds a monopoly. Considering that Skedge is an \emph{optional alternative} to CDCS, such consistent usage data (especially with 75\% of sessions being from returning users) undeniably proves its measures of usability and user satisfaction relative to CDCS.

@@

Figure \ref{fig:searchtypes} validates the hypothesis made in Section 2.3 regarding the search interface design, demonstrating that the \emph{department code} and \emph{course number} search modes dominate user searches, accounting for over 85\% of all search types on Skedge. Bringing the nondistinct field for this in CDCS into prominence is thus a design choice that accomodates empirical behavior.

@@

  This already is evidence of an effective search mechanism, but of those empty searches, the vast majority contain typos or are for courses that simply don't exist. The other significant portion of searches include instructor names without the requisite ``{\tt taught by}'' for SkedgeQL to pick it up as an instructor search, and this issue could easily be fixed.

@@

Making the case for usability superior to CDCS's in the space of mobile support is easy, given that CDCS does not have a mobile-responsive website. But to illustrate the importance of ensuring good user experience on mobile devices, note that 11.54\% of all Skedge sessions are from mobile devices, which is very substantial. On average, 2.74 pages viewed are viewed and 2 minutes are spent per session on mobile devices.

@@

  \caption[Course blocks in Skedge]{Hovering over a course block in Skedge's side-schedule shows at-a-glance information about it; clicking on it navigates the user to that course}
  \label{fig:courseblocks}

### 2016-04-14 design.tex
More splitup
submitted

From its very inception, Skedge's functionality and visual design were driven by the shortcomings of CDCS. Skedge is built \emph{bottom-up}, not \emph{top-down}---every aspect of the application was either made as a reaction to a particular grievance in CDCS or as the natural evolution of an existing feature. Skedge is thus rooted in \emph{usability} derived from real need, not mere conjecture along the question ``what could students want?''. Its success with students, shown in Chapter 4, demonstrates that this usability extends beyond my own standard and can fulfill the various discovered use-cases of students in general.

@@

\input{design-modernity}
\clearpage

@@

\input{design-usability}
\clearpage

@@

\input{design-search}
\clearpage

@@

\input{design-social}

>>>>

From its very inception, Skedge's functionality and visual design were driven by the shortcomings of CDCS. Skedge is built \emph{bottom-up}, not \emph{top-down}---every aspect of the application was either made as a reaction to a particular grievance in CDCS or as the natural evolution of an existing Skedge feature. Skedge is thus rooted in \emph{usability} derived from real need, not mere conjecture along the question ``what could students want?''. Its success with students, shown in Chapter 4, demonstrates that this usability extends beyond my own standards and preferences and can fulfill the various discovered use-cases of students in general.

@@

\input{design/modernity}
\clearpage

@@

\input{design/usability}
\clearpage

@@

\input{design/search}
\clearpage

@@

\input{design/social}

### 2016-04-14 design/modernity.tex
Slight mods
submitted

CDCS is an old system, relatively speaking, and its development on user-facing features has been almost entirely stagnant. It launched in 2009, seven years ago, and has hardly changed since. Figure \ref{fig:cdcs2010} shows CDCS in July 2010, which, besides the addition of a few search fields, is identical to its current version. Yet, since its introduction in 2009, we have seen the rise of mobile devices into ubiquity, a boom in ``hacker culture'' and public APIs, and the capability for standalone web applications to be as sophisticated and dynamic as desktop-class applications without the aid of browser extensions. With this in mind, Skedge brings course scheduling to the modern era.

\begin{figure}[ht]
  \centering
    \fbox{
      \includegraphics[width=10cm]{images/cdcs/2010}
    }
  \caption[CDCS in July 2, 2010]{CDCS in July 2, 2010, virtually unchanged from today, courtesy of \emph{Archive.org}}
  \label{fig:cdcs2010}
\end{figure}

@@

Moreover, since no major mobile browser currently supports browser extensions (and if one did, the extensions themselves would most likely need to be re-architected), CDCS on a mobile device loses all scheduling functionality, unlike Skedge.

\begin{figure}[ht]
  \centering
  \vspace{10pt}
  \begin{tabular}{c c}
    \begin{subfigure}[h]{4.5cm}
      \centering
      \includegraphics[width=1.00\textwidth]{images/cdcs/mobile}
      \caption{CDCS} \label{fig:cdcs-mobile}
    \end{subfigure}
    \begin{subfigure}[h]{4.4cm}
      \centering
      \includegraphics[width=1.00\textwidth]{images/skedge/mobile}
      \caption{Skedge} \label{fig:sk-mobile}
    \end{subfigure}
  \end{tabular}
  \caption{CDCS and Skedge running on a mobile device}
\end{figure}

@@

With the increasing number of attendees at University of Rochester's hackathons, it is clear that the University's ``hacker culture'' is growing---more students are collaborating to build side-projects that integrate resources and servies often benefitting the student community. Open-source and open-information services greatly help to foster such innovation, and having public APIs is essential toward this end \cite{Milberry}\cite{hackers}.

Skedge provides a public JSON API at the root URL \url{http://www.skedgeur.com/api/}, made at the request of a student that was interested in using its course data, and the API has already been used in projects by several other student groups. The endpoints included as of now are \url{/api/courses?q=query} (Skedge's query language---described in detail in section 2.3---is supported here) and \url{/api/departments} for a list of department names and codes.

@@

As explained in the introduction, Skedge offers users a course schedule right in the page, unlike CDCS which requires the Better CDCS browser extension for this functionality. Having a schedule native in the application has several advantages:

\begin{enumerate}
\item Besides some CDCS users possibly not even knowing about Better CDCS, not requiring a browser extension provides for a faster and more seamless user onboarding, especially when building schedules on public computers when extensions can't always be installed.

\item Skedge was designed to have a schedule, whereas Better CDCS works within an interface that wasn't. CDCS suffers lower usability as a result, requiring toggle between results and schedules. Skedge, conversely, immediately projects a course into the user's schedule.

>>>>

CDCS is an old system, relatively speaking, and its development on user-facing features has been almost entirely stagnant. It launched in 2009, seven years ago, and has hardly changed since. Figure \ref{fig:cdcs2010} shows CDCS in July 2010, which, besides the addition of a few search fields, is identical to its current version. Yet, since its introduction in 2009, we have seen the rise of mobile devices into ubiquity, a boom in ``hacker culture'' and public APIs, and the capability for standalone web applications to be as sophisticated and dynamic as desktop-class applications without the aid of browser extensions. To this end, Skedge brings course scheduling to the modern era.

\begin{figure}[ht]
  \centering
    \fbox{
      \includegraphics[width=10cm]{images/cdcs/2010}
    }
  \caption[CDCS in July 2, 2010]{CDCS in July 2, 2010, virtually unchanged from today, courtesy of \emph{Archive.org}\cite{archive}}
  \label{fig:cdcs2010}
\end{figure}

@@

Moreover, since no major mobile browser currently supports browser extensions (and if one did, the extensions themselves would most likely need to be re-architected), CDCS on a mobile device loses all scheduling functionality, unlike Skedge, which supports it.

\begin{figure}[ht]
  \centering
  \vspace{10pt}
  \begin{tabular}{c c}
    \begin{subfigure}[h]{4.8cm}
      \centering
      \includegraphics[width=1.00\textwidth]{images/cdcs/mobile}
      \caption{CDCS} \label{fig:cdcs-mobile}
    \end{subfigure}
    \begin{subfigure}[h]{4.7cm}
      \centering
      \includegraphics[width=1.00\textwidth]{images/skedge/mobile}
      \caption{Skedge} \label{fig:sk-mobile}
    \end{subfigure}
  \end{tabular}
  \caption{CDCS and Skedge on a mobile browser}
\end{figure}

@@

With the increasing number of attendees at University of Rochester's hackathons, it is clear that the University's ``hacker culture'' is growing---more students are collaborating to build side-projects that integrate resources and services often benefitting the student community. Open-source and open-information services greatly help to foster such innovation, and having public APIs is essential toward this end \cite{Milberry} \cite{hackers}.

Skedge provides a public JSON API at the root URL \url{http://www.skedgeur.com/api/}, originally made at the request of a student that was interested in using course data, and the API has since been used in projects by several other student groups. The endpoints included as of now are \url{/api/courses?q=query} (Skedge's query language---described in detail in section 2.3---is supported here) and \url{/api/departments} for a list of department names and codes.

@@

Skedge offers users a course schedule right in the page, unlike CDCS which requires the Better CDCS browser extension. Having a native schedule has several advantages:

\begin{enumerate}
\item Besides some CDCS users possibly not even knowing about Better CDCS, not requiring a browser extension provides for a faster and more seamless user onboarding, especially when building schedules on public computers where extensions can't always be installed.

\item Skedge was designed with a native schedule in mind, whereas Better CDCS works within an interface that wasn't. CDCS suffers from lower usability as a result, requiring the user to toggle between results and the schedule. Skedge, conversely, provides immediate visual feedback on how courses fit∫ into the user's schedule when the cursor is hovered over them.

### 2016-04-14 design/search.tex
submitted

In the spirit of this chapter's theme, Skedge's search method was, again, primarily designed as a reaction to CDCS's. CDCS's search method is a 15-field form (see Figure \ref{fig:cdcs-search}), of which I only found myself regularly using two (one of them being the \emph{required} ``Year/Term'' field). This prompted me to closely examine the fields to a) determine redundancies between them and b) find how to make some of the valuable filter fields more usable. From this came Skedge's unified, natural language based search method, which I call the Skedge Query Language, or \emph{SkedgeQL}. Figure \ref{fig:sk-search} is a list of example possible searches shown to users as they type.

\begin{figure}[ht]
  \centering
  \vspace{5pt}
  \begin{tabular}{c c}
    \begin{subfigure}[w]{4.5cm}
      \centering
      \fbox{
        \includegraphics[width=3.3cm]{images/cdcs/search}
      }
      \caption{Form-based search in CDCS} \label{fig:cdcs-search}
    \end{subfigure}
    \hspace{5pt}
    \begin{subfigure}[w]{8.5cm}
      \centering
      \fbox{
        \includegraphics[width=1.00\textwidth]{images/skedge/search}
      }
      \caption{Examples of SkedgeQL} \label{fig:sk-search}
    \end{subfigure}
  \end{tabular}
  \caption{Two philosophies of search method}
\end{figure}

@@

    The main advantage of Skedge's search system is in its reduction of 15 fields to a single one. I hypothesize that since the vast majority of searches fit the department code / course number pattern (this hypothesis later supported in Chapter 4), the time required to force the user to select a term and then find the correct box to enter this query is wasted.

@@

    Another notable disadvantage of SkedgeQL is having to know it, and it having a possibly steep learning curve for some users. However, the argument will be made in Chapter 4 that simply by using the site over time, users will self-learn Skedge. One of the ways this is possible is thanks to Skedge's search system being \emph{multi-purposed}. Hyperlinks around the site---instructor names (which link to ``{\tt taught by <name>}''), course references, or schedule course blocks for example---all direct the user to an ordinary search results page with the search field populated, signalling to the user that such a search is possible.

@@

I have identified \emph{three} use-cases for course searching (the existence of and distinction between these cases will be demonstrated with collected usage data in Chapter 4). The three cases are \textbf{requirements}, \textbf{electives}, and \textbf{peer recommendations}. SkedgeQL and other application features offer substantial improvements over CDCS for each of these cases.

@@

  These are courses that are required for a student to complete their degree, and are typically searched for directly. The functionality required here is simple and is mostly satisfied by CDCS, but Skedge offers the following improvements to the process:

@@

  Elective courses can either be courses that are required for a major or minor, but are chosen by the student, or they can be courses of particular interest to students who have the flexibility to take them. For this use-case, Skedge offers search, filter, and sorting features that substantially aid users in browsing through courses and discovering ones they might want to take. Note that none of the following features are supported by CDCS.

>>>>

In the spirit of this chapter's theme, Skedge's search method was, again, primarily designed as a reaction to CDCS's. CDCS's search method is a 15-field form (see Figure \ref{fig:cdcs-search}), of which I only found myself regularly using two (one of them being the \emph{required} ``Year/Term'' field). This prompted me to closely examine the fields to a) determine redundancies between them and b) find how to make some of the valuable filter fields more usable. From this came Skedge's unified, natural language based search method, which I call the Skedge Query Language, or \emph{SkedgeQL}. Figure \ref{fig:sk-search} is a list of example possible searches shown to users as they type into the searchbox.

\begin{figure}[ht]
  \centering
  \vspace{5pt}
  \begin{tabular}{c c}
    \begin{subfigure}[w]{4.5cm}
      \centering
      \fbox{
        \includegraphics[width=3.3cm]{images/cdcs/search}
      }
      \caption{Form-based search in CDCS} \label{fig:cdcs-search}
    \end{subfigure}
    \hspace{5pt}
    \begin{subfigure}[w]{8.5cm}
      \centering
      \fbox{
        \includegraphics[width=1.00\textwidth]{images/skedge/search}
      }
      \caption{Examples of SkedgeQL} \label{fig:sk-search}
    \end{subfigure}
  \end{tabular}
  \caption{Two philosophies of search}
\end{figure}

@@

    The main advantage of Skedge's search system is in its reduction of 15 fields to a single one. I hypothesize that since the vast majority of searches fit the department code / course number pattern (e.g. ``{\tt csc 171}''---this hypothesis later supported in Chapter 4), the time required to force the user to select a term and then find the correct box to enter this query is wasted.

@@

    Another notable disadvantage of SkedgeQL is having to know it, and it having a possibly steep learning curve for some users. However, the argument will be made in Chapter 4 that simply by using the site over time, users will self-learn SkedgeQL. One of the ways this is possible is thanks to Skedge's search system being \emph{multi-purposed}. Hyperlinks around the site---instructor names (which link to ``{\tt taught by <name>}''), course references, or schedule course blocks for example---all direct the user to an ordinary search results page with the search field populated, signalling to the user that such a search is possible.

@@

I have identified \emph{three} use-cases for course searching (the existence of and distinction between these cases will be demonstrated with collected usage data in Chapter 4). The types of courses user look for are \textbf{requirements}, \textbf{electives}, and \textbf{peer recommendations} (associated with direct, exploratory, and peer-guided search, respectively). SkedgeQL and other application features offer substantial improvements over CDCS for each of these cases.

@@

  This is for are courses that are required for a student to complete their degree, and are typically searched for directly. The functionality required here is simple and is mostly satisfied by CDCS, but Skedge offers the following improvements to the process:

@@

  Elective courses can either be courses that are required for a major or minor, but are chosen by the student, or they can be courses of particular interest to students who have the flexibility to take them,. For this use-case, Skedge offers search, filter, and sorting features that substantially aid users in browsing through courses and discovering ones they might want to take. Note that none of the following features are supported by CDCS.

### 2016-04-14 design/social.tex
Tech
submitted

    \begin{subfigure}[w]{7cm}
      \centering
      \includegraphics[width=6.1cm]{images/skedge/facebook}
      \caption{Sharing a Skedge schedule on Facebook, linked to the live and dynamic schedule share page on Skedge} \label{fig:sk-fb}
    \end{subfigure}

@@

    As mentioned above, the schedules generated with Better CDCS cannot be shared publicly, and therefore students can only post a static image of their schedule at the time of posting. This excludes the possibility of edits to the schedule being reflected in the same post to their friends. Moreover, referencing courses (i.e. looking up more information on a course a friend is taking) is not very usable under this model as it requires a manual search.

@@

  Immediately after logging in, the user will see the ``schedule feed,'' the list of the user's peers' current schedules (see Figure \ref{fig:sk-feed}). This takes from a familiar paradigm of the \emph{feed} in social networks, which is routinely scrolled for updates.[1] For each schedule, the user can visualize their own schedule side-by-side (using with the same ``overlay'' mechanic explained earlier), access the Facebook profile of the schedule's owner, and see the list of courses that user likes.

>>>>

    \begin{subfigure}[w]{7cm}
      \centering
      \includegraphics[width=6.1cm]{images/skedge/facebook}
      \caption{Sharing a Skedge schedule on Facebook links to the live and dynamic schedule share page} \label{fig:sk-fb}
    \end{subfigure}

@@

    As mentioned previously, the schedules generated with Better CDCS cannot be shared publicly, and therefore students can only post a static, exported image of their schedule. This excludes the possibility of edits to the schedule being reflected to their peers in the same post. Moreover, referencing courses (i.e. looking up more information on a course a peer is taking) is not very usable under this model as it requires a manual search.

@@

  Immediately after logging in, the user will see the ``schedule feed,'' the list of the user's peers' current schedules (see Figure \ref{fig:sk-feed}). This takes from a familiar paradigm of the \emph{feed} in social networks, which is routinely scrolled for updates. For each schedule, the user can visualize their own schedule side-by-side (using with the same ``overlay'' mechanic explained earlier), access the Facebook profile of the schedule's owner, and see the list of courses that user likes.

### 2016-04-14 design/usability.tex
little updates
submitted

Collapsing subsections within courses can result in massive improvements in filtering the data most relevant to the user. For instance, the search {\tt csc} for Fall 2016 on CDCS results in 147 ``section boxes'', while Skedge only shows 45 ``\emph{course} boxes'', with subsections collapsed within their respective course. This triage reduces the data (noise, more correctly) displayed by \textbf{$\sim$70\%}, and is even higher for departments with more abundant labs and workshops, such as Physics (Skedge: 35 vs. CDCS: 226, \textbf{an $\sim$85\% reduction}), or Chemistry (Skedge: 25 vs. CDCS: 171, \textbf{an $\sim$86\% reduction}).

Skedge can reduce the amount of data to scroll through---and thus the time taken to do so---by six- or seven-fold (and possibly more, counting the attention users otherwise have to pay to distinguish course from subsections), so this design decision has a large usability payoff.

@@

\begin{figure}[ht]
  \centering
  \vspace{2pt}
  \includegraphics[width=14cm]{images/skedge/export}
  \caption{Exporting to Google Calendar (note the other export options on the top right)} \label{fig:sk-export}
\end{figure}

>>>>

Collapsing subsections within courses can result in massive improvements in filtering the data most relevant to the user. For instance, the search ``{\tt csc}'' for Fall 2016 on CDCS results in 147 ``section boxes'', while Skedge only shows 45 ``\emph{course} boxes'', with subsections collapsed within their respective course. This triage reduces the data (noise, more correctly) displayed by \textbf{$\sim$70\%}, and is even higher for departments with more abundant labs and workshops, such as Physics (Skedge: 35 vs. CDCS: 226, \textbf{an $\sim$85\% reduction}), or Chemistry (Skedge: 25 vs. CDCS: 171, \textbf{an $\sim$86\% reduction}).

Skedge can reduce the amount of data to scroll through---and thus the time taken to do so---by six- or seven-fold (and possibly more, counting the attention users otherwise have to pay to distinguish courses from subsections), so this design decision has a large usability payoff.

@@

\begin{figure}[ht]
  \centering
  \vspace{2pt}
  \includegraphics[width=14cm]{images/skedge/export}
  \caption[Skedge exporting options]{Exporting to Google Calendar (note the other export options on the top right)} \label{fig:sk-export}
\end{figure}

### 2016-04-14 introduction.tex
submitted

\vspace{-20pt}

I am proud to finally, officially introduce the world to \emph{Skedge}, a website I have been developing since December 2013. It began as a weekend project after I spent too much time trying to find interesting elective classes, and it wasn't until I briefly presented it at a RocHack ``Hacker Night'' when I realized the potential it had to help others as well. I never expected the subject of my senior thesis to be something so close to my heart, so I am grateful for such a fun and motivating opportunity to continue work on Skedge, and, selfishly, for an outlet of recognition.

\section{Overview of CDCS}

@@

\section{Overview of Skedge}

@@

\noindent
This paper is organized in two parts, with a brief intermission in between. First, I will explain many of Skedge's design decisions in relation to what I call my ``grievances'' with CDCS in its current state. The second part will look at collected data, extrapolating how students use Skedge and using several novel metrics to measure its usability and effectiveness as a platform.

>>>>

\onehalfspacing
{\small
I am proud to finally and officially introduce \emph{Skedge}, a website I have been developing since December 2013. Skedge began as a weekend project after far too much time was spent trying to find interesting elective classes, and it wasn't until I briefly presented it at a RocHack ``Hacker Night'' when I realized the potential it had to help others as well. 
I never expected the subject of my senior thesis to be something so close to my heart, so I am grateful for such a fun and motivating opportunity to continue work on Skedge, and, selfishly, for an outlet of recognition.
}

\doublespacing

\section{CDCS}

@@

\section{Skedge}

@@

\noindent
This paper is organized into two parts, with a brief intermission in between. First, I will explain many of Skedge's design decisions in response to what I call ``grievances'' with CDCS. We will break for a technical overview, and the second part will look at live usage data, extrapolating how Skedge is used by students and measuring its efficacy as a platform using some novel metrics.

### 2016-04-14 tech.tex
start intro
submitted

\chapter{Technical overview}

@@

Skedge is a \emph{Ruby on Rails} web application running on a webserver called \emph{unicorn}, and routes ports 80 to it using the \emph{nginx} webserver as a reverse-proxy. Skedge uses \emph{PostgreSQL} as database, and \emph{React.js} on the front-end, a framework developed by Facebook for making responsive user interfaces. This stack is hosted on a paid (by me) virtual machine instance on DigitalOcean.

@@

At the end of each day, Skedge runs an update script that updates its course data with any new course data that exists or has changed on CDCS, endearingly called ``scraping.'' There is minimal text processing to get the data to its final form as it is on Skedge, such as title formatting into proper capitalization. Highly sensitive data such as times, location, or virtually any other field remains completely unmodified. A substianal amount of logic goes into connecting courses to their subsections (workshops, labs, etc.), as these are not explicitly connected in CDCS and sometimes requires some basic natural language processing (finding CRN numbers, title similarity, etc.) Any ambiguities in this process are reported to me and they can easily be resolved (most arise from data entry errors in CDCS).

@@

Users are purposefully light-weight on Skedge, because it aims for as little friction and overhead as possible before users can begin to search for courses and build a schedule. Thus there are no logins (besides now with Skedge Social, which is optional), and yet Skedge keeps users' schedules persistent. This is done using a single cookie on the client side that records the user ID that client is associated with, as well as a long hexadecimal string that serves as the ``user secret,'' or ``password'', which must match the one on the server for any changes to be made. This prevents users from manually changing cookie data to modify others' schedules.

The disadvantage of the cookie-based system is that each new browser or device is disjoint by default. Previously, I made it possible to retrieve and set the user secret from the user interface, allowing users to manually synchronize all of their devices' browser cookies to the same user ID and thus manage the same schedules from different devices, but this was replaced by the same feature offered through Skedge Social, as part of promoting the platform. A different way to solve the problem of better persistence while not sacrificing the trade-off of lightweight accounts, and one that doesn't require external services like Facebook will be looked into. A likely candidate at the moment is using University of Rochester NetID for schedule synchronization and improved persistence.

@@

Out of respect for privacy and a preference to \emph{not} store sensitive data, Skedge does not keep or store any Facebook data, not even users' names or lists of friends, the two pieces of data Skedge requests access to during the integration process. Skedge only stores an anonymous ID (uniquely generated for Skedge by Facebook) per account. Only a user’s name and list of friends also using Skedge are accessible via this ID, and Skedge doesn’t even store this in its database---both are used exclusively by the client, however---separate requests are made for if any given users (friends of the user) are taking certain classes, etc., and for when displaying the schedule feed.

@@

As the saying goes, ``if it's not monitored, it's not in production'' (how else would we know that it was?) \cite{monitored}.

Skedge collects usage data that is anonymous with regards to who a specific user is (because that information is not even known), but that is not anonymous with respect to a trackable user ID. Keeping track of certain properties across users is very important for the data analysis in Sections 4.2 and 4.3. This tracking is done by the open-source library \emph{Ahoy}, which also has a JavaScript library to log click events unobtrusively and efficiently (e.g. send them after the page has been loaded) using cookies. Note that this has nothing to do with advertisement tracking that is done by marketing agencies across the web---this tracking is entirely local to Skedge and Skedge does not advertise in any way.

>>>>

\chapter{Very brief technical overview}

@@

Skedge is a \emph{Ruby on Rails} web application running on a webserver called \emph{unicorn}, and routes ports 80 to it using the \emph{nginx} webserver as a reverse-proxy. Skedge uses \emph{PostgreSQL} for the database, and \emph{React.js} on the front-end, a framework developed by Facebook for responsive user interfaces. This stack is hosted on a paid (by me) virtual machine instance on DigitalOcean.

@@

At the end of each day, Skedge runs an update script that updates its course data with any new course data that exists or has changed on CDCS, endearingly called ``scraping.'' It accesses the ordinary CDCS search endpoint, but in XML format. There is minimal text processing to get the data to its final form as it is on Skedge, such as title formatting into proper capitalization. Highly important data such as times, location, or virtually any other field remains unmodified.

A substantial amount works to connect courses to their subsections (workshops, labs, etc.), as these are not explicitly connected in CDCS and sometimes requires basic natural language processing (finding CRN references, title similarity, etc.) Any ambiguities in this process are reported to me and can easily be resolved (most arise from data entry errors in CDCS).

@@

User accounts are purposefully light-weight on Skedge, because it aims for as little friction and overhead as possible before users can begin to search for courses and build a schedule. Thus there are no logins (besides now with Skedge Social, which is optional), and yet Skedge keeps users' schedules persistent. This is done using a single cookie on the client side that records the user ID that client is associated with, as well as a long hexadecimal string that serves as the ``user secret,'' or ``password'', which must match the one on the server for any changes to be made. This prevents users from manually changing cookie data to modify others' schedules.

The disadvantage of the cookie-based system is that each new browser or device is disjoint by default. Previously, I made it possible to retrieve and set the user secret from the user interface, allowing users to manually synchronize all of their devices' browser cookies to the same user ID and thus manage the same schedules from different devices, but this was replaced by the same feature offered through Skedge Social, as part of promoting the platform. A different way to solve the problem of better persistence while not sacrificing the trade-off of lightweight accounts (and one that doesn't require external services like Facebook) is being considered. A likely candidate at the moment is using University of Rochester NetID for schedule synchronization and improved persistence.

@@

Out of respect for privacy and a preference to \emph{not} store sensitive data, Skedge does not keep or store any Facebook data, not even users' names or lists of friends, the two pieces of data Skedge requests access to during the integration process. Skedge only stores an anonymous ID (uniquely generated for Skedge by Facebook) per account. Only a user’s name and list of friends also using Skedge are accessible via this ID, and Skedge doesn't store this in its database---both are retrieved exclusively by the client, and making separate requests that ask whether any given users (i.e. peers) are taking certain classes, etc., and for when displaying the schedule feed.

@@

As the saying goes, ``if it's not monitored, it's not in production'' (how else would we know that it's truly live?) \cite{monitored}.

Skedge collects usage data that is anonymous with regards to who a specific user is (because that information is not even known or stored), but that is not anonymous with respect to a trackable user ID. Keeping track of certain properties across users is very important for the data analysis in Sections 4.2 and 4.3. This tracking is done by the open-source library \emph{Ahoy}, which also has a JavaScript library to log click events unobtrusively and efficiently (e.g. send them after the page has been loaded) using cookies. Note that this has nothing to do with advertisement tracking that is done by marketing agencies across the web---this tracking is entirely local to Skedge and Skedge does not advertise in any way.
